Deep Learning

### 2015-03-30 deepreview.tex
YB's changes + GH's changes

Deep Learning techniques allow a machine composed of multiple processing layers
which define representations of the data that are more abstract for deeper layers.
These methods have dramatically improved the state-of-the-art
for speech recognition, object recognition, object detection,
predicting the activity of drug molecules, and many other tasks. Deep
learning discovers intricate structure in large datasets by using the
{\em backpropagation algorithm} to indicate how the machine should
change.  Deep {\em convolutional nets}
have brought about dramatic improvements to process
images, video, speech and audio, while {\em recurrent
  nets} have shone on sequential data such as text and
speech.
%Convolutional nets have quickly become the leading technology for image
%recognition and speech recognition, while recurrent nets are
%competitive in several natural language processing tasks, including
%language translation. 
%Combinations of recurrent and convolutional nets
%can generate good textual captions of images.

@@

% what is machine learning
Machine learning technology (ML) powers many aspects of modern
society: from web search to content filtering on social networks, and
to recommendations on e-commerce websites. It is also increasingly
present in consumer products such as cameras and smartphones. ML
systems are used to identify objects in images, transcribe speech into
text, match news items, posts, or products with users' interests, and
select relevant results of search. Increasingly, these applications
make use a new class of techniques: {\em deep learning}.

% handcrafted features
Traditional ML techniques are limited in their ability to process
natural data in raw form. For decades, constructing a pattern
recognition or ML system has required engineering efforts and domain
expertise to design a {\em feature extractor} that transforms the raw
data, such as the pixel values of an image, into a suitable {\em
  internal representation}, or feature vector, from which the learning
subsystem, often a classifier, can detect, recognize, or classify
patterns in the input.

% deep learning: learning representations
Deep learning is a set of methods that allows a machine to be fed with
raw data and to automatically discover suitable representations of the
data. The name comes from the fact that deep learning machines are
multilayer stacks of trainable modules, each of which (e.g. a neural
network layer) turns the
representation of the data in its input into a more abstract and
higher-level representation on its output in which relevant
characteristics and amplified and irrelevant variabilities are
attenuated. For example, an image comes in the form of an array of
pixel values, features in the first layer would represent the presence
or absence of edges of a particular orientation at all locations in
the image, the second layer would  detect motifs by spotting particular
arrangements of edges, regardless of small variations in the positions
of the edges.  The third layer would assemble motifs into parts of
object, and subsequent layers would to detect objects as combinations of
parts. The key aspect of deep learning is that {\em these features and
  representations are not designed by human engineers, but are
  automatically learned from data.}

% preview of deep learning results
Deep learning is making major advances in solving problems that have
resisted the best attempts of the Artificial Intelligence community
for many years. It has turned out to be very good at discovering
intricate structure in large, high-dimensional data and is therefore
applicable to many domains of science, business, and government. In
addition to beating records in image
recognition~\citep{Krizhevsky-2012-small,farabet-pami-13,tompson-nips-14,szegedy-2014}
and speech
recognition~\citep{Hinton-et-al-2012,Sainath-et-al-ICASSP2013}, it has
beaten other ML techniques at predicting the activity of potential
drug molecules~\citep{Dahl-et-al-arxiv2014,Ma-et-al-2015}, detecting new fundamental
particles~\citep{Ciodaro-et-al-2012,Melis-Higgs-boson-competition-2014},
% YB thinks that Ciodaro-et-al might be {YANNparticles} 
% classifying galaxies, ~\citep{YANNgalaxies}, % YB while waiting for bib entry, 
reconstructing brain circuits~\citep{helmstaedter-nature-2013}, and
predicting the effects of mutations in non-coding DNA on gene
expression and disease~\citep{Keung-et-al-2014,Xiong-et-al-Frey-science2015}.

Perhaps more interestingly, deep learning has produced extremely
promising results for a variety of tasks in natural language
understanding~\citep{collobert:2011b}, particularly topic
classification, %~\citep{topic-classification},  % YB while waiting for bib entry
sentiment
analysis~\citep{Glorot+al-ICML-2011-small},
question-answering~\cite{Bordes-et-al-EMNLP2014}, and language
translation~\cite{Jean-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

% supervised learning, objective function, gradient.
The most common form of machine learning, deep or not, is {\em
  supervised learning}. Imagine that we want to build a system that
can classify images as containing, say, a house, a car, a person, or a
pet. We first collect a large dataset of images of houses, cars,
persons and pets, each labeled with its category. During training, the
machine is shown an image and produces an output in the form of
a vector of scores, one for each category. We want the desired
category to have the highest score of all categories, but this is
unlikely to happen before training.  We compute an {\em objective
  function}, that measures the error (or distance) between the output
scores and the desired pattern of scores. The machine then modifies
its internal adjustable parameters so as to reduce this error. These
adjustable parameters, often called weights, are real numbers which
can be seen as ``knobs'' that define the input-output function of the
machine. In a typical deep learning system, there may be hundreds of
millions of these adjustable weights, and tens or hundreds of millions
of examples with which to train the machine.

To properly adjust the weight vector, the learning algorithm computes
a {\em gradient vector} which, for each weight, indicates by what
amount the error would increase or decrease if the weight were
increased by a tiny amount. The weight vector is then adjusted in the
opposite direction of the gradient vector.

@@

% SGD
In practice, most practitioners use a procedure called {\em Stochastic
  Gradient Descent} or SGD. It consists in showing an input, computing
the output and the error, computing the gradient, and adjusting the
weight vector right away. The process is repeated with each example in
the training set, again and again, until the average of the objective
function stops decreasing. It is called stochastic because each sample
gives a noisy estimate of the average gradient. The procedure gets
close to a good minimum very quickly in
practice~\cite{bottou-bousquet-2008-small}. Following this, the
performance of the system is measured on a different set of examples,
not seen during training, called a {\em test set}. This serves to test
the {\em generalization ability} of the machine, i.e. its ability to
produce sensible answers on new inputs that it has never seen during
training.

@@

% features: solving the selectivity-invariance dilemma
It has been known since the 1960s that linear classifiers can only
carve out their input space into very simple regions, namely
half-spaces separated by a hyperplane~\citep{Duda-Hart}.  But problems
such as image and speech recognition require the input-output function
to be insensitive to irrelevant variations of the input, such as
variations in position, orientation or illumination of an object, or
variation in pitch or accent of speech, while being very sensitive to
particular minute variations, such as the difference between a white
wolf and a samoyed (a breed of wolf-like white dog).  At the pixel
level, images of two samoyeds in different poses and in different
environments may be very different from each other, while two images
of a samoyed and a wolf in the same position and on similar
backgrounds may be quite similar to each other. A linear classifier,
or any other ``shallow'' classifiers operating on raw pixels could not
possibly distinguish the latter two, while putting the former two in
the same category. This is why shallow classifiers require a good
feature extractor that solve the {\em selectivity-invariance dilemma},
i.e. that produces representations that are {\em selective} to the
important aspects of the image, while being {\em invariant} to its
irrelevant aspects, like the pose of the animal.
%
% classical extensions of simple classifiers
% YB: I propose to mostly remove the paragraph below (but keep the Scholkopf book ref)
% There have been numerous attempts, going back to the early 1960s, to
% increase the flexibility of classifiers so as to allow them to
% separate categories from one another by more complex surfaces than
% flat hyperplanes. A popular example is the class of kernel methods~\citep{Scholkopf+Smola-2002book}, % added ref for BS
%particularly the Support Vector Machine~\citep{Boser92} which
% represents inputs by a vector whose components are similarities to
% each of the training samples computed through a so-called {\em kernel
%   function}. While these methods are good general-purpose classifiers,
% they do not perform well for tasks such as image or speech recognition
% in which the input space is very high-dimensional and the input-output
% mapping is so complex that computing similarities using a simple
% kernel function is ineffective. 
Designing feature extractors that are
sensitive to the relevant information and invariant to the irrelevant
one is what generations of computer vision and pattern recognition
researchers have spent their careers on.

% deep architectures
But what if features could be learned? That is the key idea of deep
learning.

A deep learning architecture is a multilayer stack of modules, all (or
most) of which are subject to learning, and many of which compute
non-linear input-output mappings.  Each module in the stack transforms
its input so as to increase both the selectivity and the invariance a
little bit.  With multiple non-linear layers, say 5 to 20, a system
can implement extremely intricate functions of its inputs that are
simultaneously sensitive to minute details-- distinguishing samoyeds
from wolves-- and insensitive to irrelevant variations such as the
background, the pose, the lighting, or the surrounding objects.

@@

From the earliest days of pattern
recognition~\citep{selfridge,Rosenblatt57} researchers have aimed to
replace hand-engineered features by trainable multilayer networks.
But despite its extreme simplicity, the solution did not appear until
the mid 1980s and was largely forsaken by the ML community until
recently. As it turns out, training multilayer architectures can be
done by simple stochastic gradient descent. As long as the modules are
relatively smooth functions of their inputs and of their internal
weights, one can compute gradients through the {\em backpropagation
  procedure}.

The idea that this could be done, and that it could work at all, took
a surprisingly long time to emerge. It was discovered independently by
several different groups in the 1970s and
1980s~\citep{Werbos74,Parker85,LeCun85,RHW}, and became quite popular
in the late 1980s, spurring the second wave of interest for neural
networks. But for reasons that historians of science will have to
explain, the method was the subject of scorn by the ML community
between the late 1990s and the late 2000s.

The backpropagtion procedure to compute gradient of an objective
function with respect to the weights of a multilayer stack of modules
is nothing more than a practical application of the chain rule for
derivatives. The key insight is that the derivative (or gradient) of
the objective with respect to the {\it input} of a module can be
computed by working backwards from the gradient with respect to the
{\em output} of that module (or the input of the subsequent
module), as illustrated in Figure~\ref{fig:backprop-box}.
% Basically, take a module $M$ inside the network, whose input
%is $x$, weight vector is $w$, and output is $y = M(x,w)$. If one knows
%the gradient $\deri{E}{y}$ of the objective function $E$ with respect
%to the module's output $y$, one can blindly apply chain rule to
%compute the gradient with respect to the input:
%\[
%  \deri{E}{x} = \deri{y}{x}\deri{E}{y}
%\]
%This equation works even when $x$ and $y$ are vectors, in which case
%$\deri{E}{x}$ and $\deri{E}{y}$ are gradient vectors, and
%$\deri{y}{x}$ is a {\em Jacobian matrix} of $M$ where element $(i,j)$
%indicates by how much output $j$ would increase proportionally if
%input $i$ were increased by a tiny amount $\Delta y_j =
%\left(\deri{y}{x}\right)_{ij} \Delta x_i$. 
The backpropagation equation
constitutes a recurrence formula that can be used to propagate
gradients from the top layer all the way to the bottom. Once all these
gradients are computed, one can similarly compute the gradients with
respect to the weights of the module. 
%Figure~\ref{fig:backprop-box})
%shows how this works out when the network is composed of layers of
%weight matrices interspersed with scalar non-linearities.
%% the notation in the figure should be updated.

% deep neural nets
Many applications of deep learning use feedforward neural network
architectures, illustrated in Figure~\ref{fig:backprop-box}, which
learn to map a fixed-size input (e.g., an image) to a fixed-size
output (e.g. a probability for each one of several categories). To go
from one layer to the next, a set of units compute a weighted
sum of their inputs from the previous layer and pass the result
through a non-linear function. The most common non-linear function is
the so-called {\em ReLU} (Rectified Linear Unit) which is simply the
half-wave rectifier $f(z) = \max(z,0)$. In past decades, neural nets
used smoother non-linearities, such as $\tanh(z)$, but the ReLU seems
to work better in networks with many layers. Units that are not in the
output layer are traditionally called {\em hidden units}. The first
layers can be seen as distorting the input in a non-linear way so that
categories become linearly separable by the last layer, as indicated
in figure~\ref{fig:backprop-box}.

@@

Vietnam and Iraq. This is not to say that Tuesday and Wednesday are
put in the same category, but rather that they share many semantic and
grammatical attributes, so that what the neural network learns from
one day of the week can often be generalized to the other, while
individual days remain different (for example one imagines that the
neural network also learns some features that distinguishes week-end
days and weekdays, while Saturday and Sunday might have special roles
in some religions, etc.).  Such representations are called {\em
  distributed representations} because its elements (the features) are
not mutually exclusive and there exists a rich and vast set of
possible configurations of values of these features, corresponding to
the variations seen in the observed data.  These word-vectors are thus

@@

humans, but automatically discovered by the neural network.  Vector

@@

The issue of representation lies at the heart of the debate between
the logic-inspired and the neurally-inspired paradigms for
cognition. In the logic-inspired paradigm, an instance of a symbol is
something whose only property is that it is either identical or
non-identical to other symbol instances. It has no internal structure
that is relevant to its use. By contrast, neural networks use large
activity vectors instead of symbols and parallel matrix operations
followed by non-linearities instead of formal rules of inference. The
elements of such vectors can be understood as the values of learned
attributes (or features) associated with some object.  With sufficient
data, computational power and depth, the combination of
backpropagation for computing gradients and stochastic gradient
descent for updating the weights on connections has proved to be very
good at learning appropriate distributed representations. 
%These
%representations allow a lot of parallel computation at the feature
%level so that complex tasks can be solved in a very small number of
%sequential steps.
Deep learning theory shows that deep nets can benefit from
an exponential advantage over classical learning algorithm not based on distributed 
representations~\citep{Bengio-localfailure-NIPS-2006-small,Bengio-decision-trees10},
 in two ways, both arising from the power of {\em composition}, so long as the
underlying data generating distribution has such a componential structure, i.e.,
is generated by a hierarchy of composition of factors~\citep{Bengio-Courville-Vincent-TPAMI2013}.
First, learning distributed representations enables generalization to new
combinations of the learned features (e.g. $2^n$ combinations, with $n$ features)
~\citep{Bengio-2009-book,Montufar+Morton-2014}. 
Second composing layers in a deep net brings the potential for another exponential
advantage~\citep{Montufar-et-al-NIPS2014} (exponential in the depth).

@@

%\begin{figure}[ht]
\begin{figure}[b]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{word-embeddings.png}
\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
}
\caption{Left: Illustration of word representations learned for modeling
language, non-linearly projected to 2-D for visualization using the
t-SNE algorithm~\citep{VanDerMaaten08}. Right: 2-D representation
of phrases learned by an English-to-French encoder-decoder recurrent
neural network or RNN~\citep{Cho-et-al-EMNLP2014}. One can observe that
semantically similar words or sequences of words are  mapped to
nearby representations. The distributed representations of
words are obtained by using backpropagation to jointly learn 
a representation for each word and
a function that predicts a target quantity such as the next word
in a sequence (for language modeling) or a whole sequence
of translated words (for machine translation)~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
Sutskever-et-al-NIPS2014}.
}
\label{fig:word-embeddings}
\end{minipage}
}
\end{figure}

@@

Thanks to advances in their architecture~\citep{Hochreiter+Schmidhuber-1997,ElHihi+Bengio-nips8-small} 
and ways of training them~\citep{Sutskever-thesis2012,Pascanu+al-ICML2013-small}, RNNs
have been found to be very good at predicting the next character in
text~\citep{Sutskever-et-al-ICML2011} or the next word in a
sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
more complex tasks.  After reading an English sentence one word at a
time, for example, an English ``encoder'' network can be trained so
that the final state vector of its hidden units is a good
representation of the thought expressed by the sentence.  This thought
vector can then be used as the initial hidden state of (or as extra
input to) a jointly trained French ``decoder'' network which outputs a
probability distribution for the first word of the French
translation. If a particular first word is chosen from this
distribution and provided as input to the decoder network it will then
output a probability distribution for the second word of the
translation and so on until a full stop is
chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

This rather naive way of performing machine translation has quickly become
competitive with the state-of-the-art and this raises serious doubts about
whether understanding a sentence requires anything like internal symbolic
expressions that are manipulated by binding their symbols to variables
in judiciously chosen rules of inference, as was previously thought by
many early AI researchers.  The encoder-decoder translation
networks do not work by using internal rules of inference in which
variables get bound to symbols. They just use big state vectors, big weight
matrices and scalar non-linearities to get the job done.  
Deriving conclusions using only the form of the premises (which is
what formal logic does) is a powerful
technique that people can eventually master, but the translation
results with RNNs suggest that it may not be a good model of
the way we normally reason or understand language. RNNs
that represent thoughts as big vectors are more compatible with
the view that everyday reasoning involves many simultaneous analogies that
each contribute plausibility to a 
conclusion~\citep{Lakoff+Johnson-2008,Rogers+McClelland-book2004}.

@@

%When LSTMs were first described \citep{Hochreiter+Schmidhuber-1997},
%their complexity may have been off-putting and, despite their
%impressive ability to learn very long-range temporal structure on toy
%problems, they were largely ignored until they were shown to yield
%excellent results in a cursive handwriting recognition
%competition~\citep{Graves-et-al-2009}. 
LSTM have subsequently proved
to be very effective for tasks that have sequential input.  When they
use several hidden layers, they are better than deep feedforward
networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
unlike feedforward nets they can implement an entire speech
recognition system that goes all the way from coefficents derived from
the soundwave to the sequence of characters in the transcription.
LSTMs or related forms of gated units are also what is currently used
for the encoder and decoder networks that perform so well at machine
translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

>>>>

Deep Learning techniques allow machines that are composed of multiple processing layers
to learn representations of data that become progressively more abstract in deeper layers.
These methods have dramatically improved the state-of-the-art
for speech recognition, object recognition, object detection,
predicting the activity of drug molecules, and many other tasks. Deep
learning discovers intricate structure in large datasets by using the
{\em backpropagation algorithm} to indicate how the machine should
change the weights that are used to compute the representation in each layer from
the representation in the previous layer.  Deep {\em convolutional nets}
have brought about dramatic improvements in processing
images, video, speech and audio, while {\em recurrent
  nets} have shone on sequential data such as text and
speech.
%Convolutional nets have quickly become the leading technology for image
%recognition and speech recognition, while recurrent nets are
%competitive in several natural language processing tasks, including
%language translation. 
%Combinations of recurrent and convolutional nets
%can generate good textual captions of images.

@@

% what is machine learning
Machine learning technology (ML) powers many aspects of modern
society: from web search to content filtering on social networks 
to recommendations on e-commerce websites, and it is increasingly
present in consumer products such as cameras and smartphones. ML
systems are used to identify objects in images, transcribe speech into
text, match news items, posts, or products with users' interests, and
select relevant results of search. Increasingly, these applications
make use a new class of techniques called {\em deep learning}.

% handcrafted features
Traditional ML techniques were limited in their ability to process
natural data in raw form. For decades, constructing a pattern
recognition or ML system required careful engineering and considerable
domain
expertise to design a {\em feature extractor} that transformed the raw
data, such as the pixel values of an image, into a suitable {\em
  internal representation}, or feature vector, from which the learning
subsystem, often a classifier, could detect or classify
patterns in the input.

% deep learning: learning representations
Deep learning is a set of methods that allows a machine to be fed with
raw data and to automatically discover the representations needed for
detection or classification. The name comes from the fact that deep learning machines are
multilayer stacks of trainable modules, each of which turns the
representation of the data on its input into a more abstract and
higher-level representation on its output. The output representation
amplifies aspects of the input that are important for discrimination
and suppresses irrelevant variations. An image, for example, comes in the form of an array of
pixel values and the learned features in the first layer of representation
typically represent the presence
or absence of edges at  particular orientations and locations in
the image. The second layer typically detects motifs by spotting particular
arrangements of edges, regardless of small variations in the edge positions
The third layer  may assemble motifs into larger
combinations that correspond to parts of familiar
objects, and subsequent layers would to detect objects as combinations of
these parts. The key aspect of deep learning is that these layers of
features are not designed by human engineers: They are learned from
data using a general purpose learning procedure.}

% preview of deep learning results
Deep learning is making major advances in solving problems that have
resisted the best attempts of the Artificial Intelligence community
for many years. It has turned out to be very good at discovering
intricate structure in large, high-dimensional data and is therefore
applicable to many domains of science, business, and government. In
addition to beating records in image
recognition~\citep{Krizhevsky-2012-small,farabet-pami-13,tompson-nips-14,szegedy-2014}
and speech
recognition~\citep{Hinton-et-al-2012,Sainath-et-al-ICASSP2013}, it has
beaten other ML techniques at predicting the activity of potential
drug molecules~\citep{Ma-et-al-2015}, detecting new fundamental
particles~\citep{Ciodaro-et-al-2012,Melis-Higgs-boson-competition-2014},
% YB thinks that Ciodaro-et-al might be {YANNparticles} 
% classifying galaxies, ~\citep{YANNgalaxies}, % YB while waiting for bib entry, 
reconstructing brain circuits~\citep{helmstaedter-nature-2013}, and
predicting the effects of mutations in non-coding DNA on gene
expression and disease~\citep{Keung-et-al-2014,Xiong-et-al-Frey-science2015}.

Perhaps more surprisingly, deep learning has produced extremely
promising results for a variety of tasks in natural language
understanding~\citep{collobert:2011b}, particularly topic
classification, %~\citep{topic-classification},  % YB while waiting for bib entry
sentiment
analysis~\citep{Glorot+al-ICML-2011-small},
question-answering~\cite{Bordes-et-al-EMNLP2014}, and language
translation~\cite{Jean-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

% supervised learning, objective function, gradient.
The most common form of machine learning, deep or not, is {\em
  supervised learning}. Imagine that we want to build a system that
can classify images as containing, say, a house, a car, a person, or a
pet. We first collect a large dataset of images of houses, cars,
persons and pets, each labeled with its category. During training, the
machine is shown an image and produces an output in the form of
a vector of scores, one for each category. We want the desired
category to have the highest score of all categories, but this is
unlikely to happen before training.  We compute an {\em objective
  function}, that measures the error (or distance) between the output
scores and the desired pattern of scores. The machine then modifies
its internal adjustable parameters so as to reduce this error. These
adjustable parameters, often called weights, are real numbers which
can be seen as ``knobs'' that define the input-output function of the
machine. In a typical deep learning system, there may be hundreds of
millions of these adjustable weights, and hundreds of millions
of labeled examples with which to train the machine.

To properly adjust the weight vector, the learning algorithm computes
a {\em gradient vector} which, for each weight, indicates by what
amount the error would increase or decrease if the weight were
increased by a tiny amount. The weight vector is then adjusted in the
opposite direction to the gradient vector.

@@

% SGD
In practice, most practitioners use a procedure called {\em Stochastic
  Gradient Descent} or SGD. It consists in showing the input vector
for a few examples, computing the outputs and the errors, computing
the average gradient for those examples, and adjusting the
weights. The process is repeated for many small sets of examples
from the training set until the average of the objective function stops
decreasing. It is called stochastic because each small set of examples
gives a noisy estimate of the average gradient over all examples. In
practice, this simple procedure usually finds a good set of weights
surprisingly quickly when compared with far more elaborate
optimization techniques~\cite{bottou-bousquet-2008-small}. After
training, the performance of the system is measured on a different set
of examples called a {\em test set}. This serves to test the {\em
  generalization ability} of the machine, {\it i.e.} its ability to
produce sensible answers on new inputs that it has never seen during
training.

@@

% features: solving the selectivity-invariance dilemma
It has been known since the 1960s that linear classifiers can only
carve their input space into very simple regions, namely
half-spaces separated by a hyperplane~\citep{Duda-Hart}.  But problems
such as image and speech recognition require the input-output function
to be insensitive to irrelevant variations of the input, such as
variations in position, orientation or illumination of an object, or
variation in the pitch or accent of speech, while being very sensitive to
particular minute variations, such as the difference between a white
wolf and a samoyed (a breed of wolf-like white dog).  At the pixel
level, images of two samoyeds in different poses and in different
environments may be very different from each other, while two images
of a samoyed and a wolf in the same position and on similar
backgrounds may be very similar to each other. A linear classifier,
or any other ``shallow'' classifier operating on raw pixels could not
possibly distinguish the latter two, while putting the former two in
the same category. This is why shallow classifiers require a good
feature extractor that solves the {\em selectivity-invariance dilemma},
{\it i.e.} that produces representations that are {\em selective} to the
aspects of the image that are important for discrimination while being {\em invariant} to 
irrelevant aspects like the pose of the animal.
%
% classical extensions of simple classifiers
% YB: I propose to mostly remove the paragraph below (but keep the Scholkopf book ref)
% There have been numerous attempts, going back to the early 1960s, to
% increase the flexibility of classifiers so as to allow them to
% separate categories from one another by more complex surfaces than
% flat hyperplanes. A popular example is the class of kernel methods~\citep{Scholkopf+Smola-2002book}, % added ref for BS
%particularly the Support Vector Machine~\citep{Boser92} which
% represents inputs by a vector whose components are similarities to
% each of the training samples computed through a so-called {\em kernel
%   function}. While these methods are good general-purpose classifiers,
% they do not perform well for tasks such as image or speech recognition
% in which the input space is very high-dimensional and the input-output
% mapping is so complex that computing similarities using a simple
% kernel function is ineffective. 
Designing good feature extractors traditionally required a lot of
engineering skill and domain expertise, but this can all be avoided if
good features can be learned automatically using a general purpose
learning procedure.  That is the key idea of deep
learning.

% deep architectures

A deep learning architecture is a multilayer stack of modules, all (or
most) of which are subject to learning, and many of which compute
non-linear input-output mappings.  Each module in the stack transforms
its input so as to increase both the selectivity and the invariance of
the representation. With multiple non-linear layers, say 5 to 20, a system
can implement extremely intricate functions of its inputs that are
simultaneously sensitive to minute details-- distinguishing samoyeds
from wolves-- and insensitive to large irrelevant variations such as the
background, pose, lighting, and surrounding objects.

@@

From the earliest days of pattern
recognition~\citep{selfridge,Rosenblatt57} researchers have aimed to
replace hand-engineered features by trainable multilayer networks,
but despite its extreme simplicity, the solution was not widely
understood until
the mid 1980s and was largely forsaken by the ML community until
recently. As it turns out, training multilayer architectures can be
done by simple stochastic gradient descent. As long as the modules are
relatively smooth functions of their inputs and of their internal
weights, one can compute gradients using the {\em backpropagation
  procedure}.

The idea that this could be done, and that it worked, was discovered independently by
several different groups in the 1970s and
1980s~\citep{Werbos74,Parker85,LeCun85,RHW}, and became quite popular
in the late 1980s, spurring a wave of interest for neural
networks.

The backpropagtion procedure to compute gradient of an objective
function with respect to the weights of a multilayer stack of modules
is nothing more than a practical application of the chain rule for
derivatives. The key insight is that the derivative (or gradient) of
the objective with respect to the {\it input} of a module can be
computed by working backwards from the gradient with respect to the
{\em output} of that module (or the input of the subsequent
module), as illustrated in Figure~\ref{fig:backprop-box}.
% Basically, take a module $M$ inside the network, whose input
%is $x$, weight vector is $w$, and output is $y = M(x,w)$. If one knows
%the gradient $\deri{E}{y}$ of the objective function $E$ with respect
%to the module's output $y$, one can blindly apply chain rule to
%compute the gradient with respect to the input:
%\[
%  \deri{E}{x} = \deri{y}{x}\deri{E}{y}
%\]
%This equation works even when $x$ and $y$ are vectors, in which case
%$\deri{E}{x}$ and $\deri{E}{y}$ are gradient vectors, and
%$\deri{y}{x}$ is a {\em Jacobian matrix} of $M$ where element $(i,j)$
%indicates by how much output $j$ would increase proportionally if
%input $i$ were increased by a tiny amount $\Delta y_j =
%\left(\deri{y}{x}\right)_{ij} \Delta x_i$. 
The backpropagation equation can be used repeatedly to propagate
gradients with respect to the outputs of a module from the top layer all the way to the bottom. Once these
gradients have  computed, it is straightforward to compute the gradients with
respect to the weights of each module. 
%Figure~\ref{fig:backprop-box})
%shows how this works out when the network is composed of layers of
%weight matrices interspersed with scalar non-linearities.
%% the notation in the figure should be updated.

% deep neural nets
Many applications of deep learning use feedforward neural network
architectures, illustrated in Figure~\ref{fig:backprop-box}, which
learn to map a fixed-size input ({\it e.g.} an image) to a fixed-size
output ({\it e.g.} a probability for each one of several categories). To go
from one layer to the next, a set of units compute a weighted
sum of their inputs from the previous layer and pass the result
through a non-linear function. The most common non-linear function is
the so-called {\em ReLU} (Rectified Linear Unit) which is simply the
half-wave rectifier $f(z) = \max(z,0)$. In past decades, neural nets
used smoother non-linearities, such as $\tanh(z)$ or $1/(1+\exp(-z))$ but the ReLU 
typically learns much faster in networks with many layers. Units that are not in the
output layer are traditionally called {\em hidden units}. The early
layers can be seen as distorting the input in a non-linear way so that
categories become linearly separable by the last layer, as indicated
in figure~\ref{fig:backprop-box}.

@@

Vietnam and Iraq. 
%% This is not to say that Tuesday and Wednesday are
%% put in the same category, but rather that they share many semantic and
%% grammatical attributes, so that what the neural network learns from
%% one day of the week can often be generalized to the other, while
%% individual days remain different (for example one imagines that the
%% neural network also learns some features that distinguishes week-end
%% days and weekdays, while Saturday and Sunday might have special roles
%% in some religions, etc.).  
Such representations are called {\em
  distributed representations} because their elements (the features) are
not mutually exclusive and their exponentially many configurations
%% exists a rich and vast set of
%% possible configurations of values of these features, 
correspond to
the variations seen in the observed data.  These word-vectors are
composed of

@@

experts, but automatically discovered by the neural network.  Vector

@@

The issue of representation lies at the heart of the debate between
the logic-inspired and the neurally-inspired paradigms for
cognition. In the logic-inspired paradigm, an instance of a symbol is
something whose only property is that it is either identical or
non-identical to other symbol instances. It has no internal structure
that is relevant to its use and  to reason with symbols, they must be
bound to the variables in judiciously chosen rules of inference.  By
contrast, neural networks just use big activity vectors, big weight
matrices and scalar non-linearities to perform the type of fast
``intuitive'' inference that underpins effortless commonsense
reasoning.

%representations allow a lot of parallel computation at the feature
%level so that complex tasks can be solved in a very small number of
%sequential steps.
Deep learning theory shows that deep nets can benefit from
an exponential advantage over classical learning algorithm not based on distributed 
representations~\citep{Bengio-localfailure-NIPS-2006-small},
 in two ways, both arising from the power of {\em composition}, so long as the
underlying data generating distribution has such a componential structure, i.e.,
is generated by a hierarchy of composition of factors~\citep{Bengio-Courville-Vincent-TPAMI2013}.
First, learning distributed representations enables generalization to new
combinations of the learned features (e.g. $2^n$ combinations, with $n$ features)
~\citep{Bengio-2009-book,Montufar+Morton-2014}. 
Second composing layers in a deep net brings the potential for another exponential
advantage~\citep{Montufar-et-al-NIPS2014} (exponential in the depth).

@@

%\begin{figure}[ht]
\begin{figure}[b]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{word-embeddings.png}
\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
}
\caption{Left: Illustration of word representations learned for modeling
language, non-linearly projected to 2-D for visualization using the
t-SNE algorithm~\citep{VanDerMaaten08}. Right: 2-D representation
of phrases learned by an English-to-French encoder-decoder recurrent
neural network or RNN~\citep{Bahdanau-et-al-ICLR2015}. One can observe that
semantically similar words or sequences of words are  mapped to
nearby representations. The distributed representations of
words are obtained by using backpropagation to jointly learn 
a representation for each word and
a function that predicts a target quantity such as the next word
in a sequence (for language modeling) or a whole sequence
of translated words (for machine translation)~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
}
\label{fig:word-embeddings}
\end{minipage}
}
\end{figure}

@@

Thanks to advances in their architecture~\citep{Hochreiter+Schmidhuber-1997,ElHihi+Bengio-nips8-small} 
and ways of training them~\citep{Sutskever-thesis2012,Pascanu+al-ICML2013-small}, RNNs
have been found to be very good at predicting the next character in
text~\citep{Sutskever-et-al-ICML2011} or the next word in a
sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
more complex tasks.  After reading an English sentence one word at a
time, for example, an English ``encoder'' network can be trained so
that the final state vector of its hidden units is a good
representation of the thought expressed by the sentence.  This thought
vector can then be used as the initial hidden state of (or as extra
input to) a jointly trained French ``decoder'' network which outputs a
probability distribution for the first word of the French
translation. If a particular first word is chosen from this
distribution and provided as input to the decoder network it will then
output a probability distribution for the second word of the
translation and so on until a full stop is
chosen~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

This rather naive way of performing machine translation has quickly become
competitive with the state-of-the-art and this raises serious doubts about
whether understanding a sentence requires anything like internal symbolic
expressions that are manipulated by using inference rules.
%%binding their symbols to variables
%%in judiciously chosen rules of inference, as was previously thought by
%%many early AI researchers. It is 
%%The encoder-decoder translation
%%networks do not work by using internal rules of inference in which
%%variables get bound to symbols. They just use big state vectors, big weight
%%matrices and scalar non-linearities to get the job done. 
%%Deriving conclusions using only the form of the premises (which is
%%what formal logic does) is a powerful
%%technique that people can eventually master, but the translation
%%results with RNNs suggest that it may not be a good model of
%%the way we normally reason or understand language.
%%RNNs that represent thoughts as big vectors are 
It is more compatible with
the view that everyday reasoning involves many simultaneous analogies that
each contribute plausibility to a 
conclusion~\citep{Lakoff+Johnson-2008,Rogers+McClelland-book2004}.

@@

%When LSTMs were first described \citep{Hochreiter+Schmidhuber-1997},
%their complexity may have been off-putting and, despite their
%impressive ability to learn very long-range temporal structure on toy
%problems, they were largely ignored until they were shown to yield
%excellent results in a cursive handwriting recognition
%competition~\citep{Graves-et-al-2009}. 
LSTM have subsequently proved
to be very effective for tasks that have sequential input.  When they
use several hidden layers, they are better than deep feedforward
networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
unlike feedforward nets they can implement an entire speech
recognition system that goes all the way from coefficents derived from
the soundwave to the sequence of characters in the transcription.
LSTMs or related forms of gated units are also what is currently used
for the encoder and decoder networks that perform so well at machine
translation~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.

Over the last year, several authors have made different proposals to
augment RNNs with a memory module. 
%% This includes simplified
%% version of the LSTM
%% circuit~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014}.
%% The results suggest that not all the elements of the LSTM architecture
%% are required.
Proposals include the ``Neural Turing Machine'' in which the
network is augmented by a ``tape-like'' memory that the RNN
can chose to read from or write to~\citep{Graves-et-al-arxiv2014}, and
``Memory Networks'' in which a regular network is augmented by a
kind of associative memory~\citep{weston-memorynet-2014}. Memory
networks have yielded excellent performance on standard
question-answering benchmarks. The memory is used to remember the
story about which the network is asked to answer questions.
